---
title: "StatisticalAnalysis"
author: "anonymous"
date: "2025-12-09"
output: html_document
---

```{r}
#Required packages
if(!require(readxl)){install.packages("readxl"); library(readxl)}
if(!require(correlation)){install.packages("correlation"); library(correlation)}
if(!require(corrplot)){install.packages("corrplot"); library(corrplot)}
if(!require(car)){install.packages("car"); library(car)}
if(!require(dplyr)){install.packages("dplyr"); library(dplyr)}
if(!require(writexl)){install.packages("writexl"); library(writexl)}
if(!require(mgcv)){install.packages("mgcv"); library(mgcv)}
if(!require(rpart)){install.packages("rpart"); library(rpart)}
if(!require(ggplot2)){install.packages("ggplot2"); library(ggplot2)}
if(!require(ggeffects)){install.packages("ggeffects"); library(ggeffects)}
if(!require(gridExtra)){install.packages("gridExtra"); library(gridExtra)}
if(!require(patchwork)){install.packages("patchwork"); library(patchwork)}
if(!require(lme4)){install.packages("lme4"); library(lme4)}
if(!require(Matrix)){install.packages("Matrix"); library(Matrix)}
if(!require(lmtest)){install.packages("lmtest"); library(lmtest)}
if(!require(gamm4)){install.packages("gamm4"); library(gamm4)}
if(!require(sjPlot)){install.packages("sjPlot"); library(sjPlot)}
if(!require(sjmisc)){install.packages("sjmisc"); library(sjmisc)}
if(!require(sjlabelled)){install.packages("sjlabelled"); library(sjlabelled)}
if(!require(performance)){install.packages("performance"); library(performance)}
if(!require(flexmix)){install.packages("flexmix"); library(flexmix)}
if(!require(glmmTMB)){install.packages("glmmTMB"); library(glmmTMB)}
if(!require(DHARMa)){install.packages("DHARMa"); library(DHARMa)}
```

### Import Dataset

```{r}
#Import data
week_kuds <- read.csv("dataset_new1.csv", sep= ";")

#create week variable without the year
names(week_kuds)[names(week_kuds) == "Week"] <- "WeekYear"
week_kuds$Week <- substr(week_kuds$WeekYear, 1, 2)
#create year variable without the week
week_kuds$Year <- substr(week_kuds$WeekYear, 4, 7)

week_kuds$Spawn <- with(week_kuds, ifelse((SpawnSeason == "SS" & Week %in% c("11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23", "24", "25", "26", "27", "28", "29", "30", "31", "32", "33", "34", "35", "36", "37", "38", "39", "40")) |
                          (SpawnSeason == "A" & Week %in% c("41", "42", "43", "44", "45", "46", "47", "48", "49", "50")) |
                          (SpawnSeason == "W" & Week %in% c("51", "52", "53", "54", "01", "02", "03", "04", "05", "06", "07", "08", "09", "10")),
                          "yes", "no"))
week_kuds$SpawnSeason <- as.factor(week_kuds$SpawnSeason)


week_kuds$File <- as.factor(week_kuds$File)
week_kuds$Species <- as.factor(week_kuds$Species)
week_kuds$Transmitter <- as.factor(week_kuds$Transmitter)
week_kuds$KUD50 <- as.numeric(week_kuds$KUD50)
week_kuds$KUD95 <- as.numeric(week_kuds$KUD95)
week_kuds$Habitat <- as.factor(week_kuds$Habitat)
week_kuds$Migration <- as.factor(week_kuds$Migration)
week_kuds$ComImport <- as.factor(week_kuds$ComImport)
week_kuds$Length_cm <- as.numeric(week_kuds$Length_cm)
week_kuds$LengthStd <- as.numeric(week_kuds$LengthStd)
week_kuds$BodyMass <- as.numeric(week_kuds$BodyMass)
week_kuds$BodyMassStd <- as.numeric(week_kuds$BodyMassStd)
week_kuds$Longevity <- as.numeric(week_kuds$Longevity)
week_kuds$Vulnerability <- as.numeric(week_kuds$Vulnerability)
week_kuds$Troph <- as.numeric(week_kuds$Troph)
week_kuds$ReceiverDensity <- as.numeric(week_kuds$ReceiverDensity)
week_kuds$MonitArea_km2 <- as.numeric(week_kuds$MonitArea_km2)
week_kuds$MCP_km2 <- as.numeric(week_kuds$MCP_km2)
week_kuds$NReceivers <- as.numeric(week_kuds$NReceivers)
week_kuds$MaxDistReceivers <- as.numeric(week_kuds$MaxDistReceivers)
week_kuds$MaxLength <- as.numeric(week_kuds$MaxLength)
week_kuds$MaxBodyMass <- as.numeric(week_kuds$MaxBodyMass)
week_kuds$a <- as.numeric(week_kuds$a)
week_kuds$b <- as.numeric(week_kuds$b)
week_kuds$Week <- as.factor(week_kuds$Week)
week_kuds$Year <- as.factor(week_kuds$Year)
week_kuds$Spawn <- as.factor(week_kuds$Spawn)
week_kuds$SpawnSeason <- as.factor(week_kuds$SpawnSeason)
```


```{r}
#Summary of the variables of the dataset
summary(week_kuds)
```

```{r}
#Code to generate information for Table S1

file_summary <- week_kuds %>%
  group_by(File) %>%
  summarise(
    n_transmitters = n_distinct(Transmitter),
    n_KUD95 = sum(!is.na(KUD95))
  )

#Code to generate information for Table S3

species_summary <- week_kuds %>%
  group_by(Species) %>%
  summarise(
    mean_length = mean(Length_cm, na.rm = TRUE),
    mean_body_mass = mean(BodyMass, na.rm = TRUE),
    mean_body_mass_std = mean(BodyMassStd, na.rm = TRUE),
    mean_length_std = mean(LengthStd, na.rm = TRUE)
  )

#Code to generate information for Table S5

mean <- week_kuds %>%
  group_by(Species) %>%
  summarise(
    n = n(),
    mean_KUD95 = round(mean(KUD95, na.rm = TRUE),3),
    sd_KUD95   = round(sd(KUD95, na.rm = TRUE),3),
    min_KUD95  = round(min(KUD95, na.rm = TRUE),3),
    max_KUD95  = round(max(KUD95, na.rm = TRUE),3),
    mean_KUD50 = round(mean(KUD50, na.rm = TRUE),3),
    sd_KUD50   = round(sd(KUD50, na.rm = TRUE),3),
    min_KUD50  = round(min(KUD50, na.rm = TRUE),3),
    max_KUD50  = round(max(KUD50, na.rm = TRUE),3)
  )

write_xlsx(mean, "meanKUD_speciessummary.xlsx")

#Code to generate information for Figure 2
mean <- mean[order(mean$mean_KUD95, decreasing = FALSE), ]
nsp <- nrow(mean)
par(mar=c(5,6,4,10)) 
# Inferior axis: KUD95
plot(mean$mean_KUD95, 1:nsp - 0.2,  
     xlim=c(0, 18),
     ylim=c(0.5, nsp + 0.5),
     yaxt="n",
     xlab="",
     ylab="Species",
     cex.lab = 1.3,
     pch=20,
     col="deepskyblue4")
axis(1, at=seq(0,18,5))
mtext(expression(KUD95 ~ (km^2)), side=1, line=2.5, col="deepskyblue4", font=2, cex=1.3)
axis(2, at=1:nsp, labels=mean$Species, las=2, cex.axis=0.8)

# KUD95 bars
for(i in 1:nsp){
  y95 <- i - 0.2
  segments(mean$mean_KUD95[i] - mean$sd_KUD95[i], y95,
           mean$mean_KUD95[i] + mean$sd_KUD95[i], y95, lwd=2, col="deepskyblue4")
  segments(mean$min_KUD95[i], y95,
           mean$max_KUD95[i], y95, lwd=1, lty=3, col="deepskyblue4")
}

# Superior axis: KUD50 
par(new=TRUE)
plot(mean$mean_KUD50, 1:nsp + 0.2,  
     xlim=c(0, 4),
     ylim=c(0.5, nsp + 0.5),
     yaxt="n", xaxt="n", ylab="", xlab="",
     pch=20, bty="n", col="forestgreen")

axis(3, at=seq(0,4,1))
mtext(expression(KUD50 ~ (km^2)), side=3, line=1.6, col="forestgreen", font=2, cex=1.3)

# KUD50 bars
for(i in 1:nsp){
  y50 <- i + 0.2
  segments(mean$mean_KUD50[i] - mean$sd_KUD50[i], y50,
           mean$mean_KUD50[i] + mean$sd_KUD50[i], y50, lwd=2, col="forestgreen")
  segments(mean$min_KUD50[i], y50,
           mean$max_KUD50[i], y50, lwd=1, lty=3, col="forestgreen")
}

#legend
abbr <- mean$Species
full <- c("Serranus atricauda", "Sparisoma cretense", "Epinephelus marginatus", "Labrus bergylta", "Sciaena umbra", "Xyrichtys novacula", "Serranus scriba", "Scorpaena porcus", "Scorpaena scrofa", "Pagrus pagrus", "Serranus cabrilla", "Solea senegalensis", "Diplodus vulgaris", "Seriola rivoliana", "Diplodus sargus", "Umbrina cirrosa", "Gadus morhua", "Spondyliosoma cantharus", "Pagellus erythrinus", "Lithognathus mormyrus", "Pseudocaranx dentex", "Dentex dentex", "Sphyraena viridensis", "Dactylopterus volitans", "Sparus aurata", "Diplodus cervinus", "Dicentrarchus labrax", "Seriola dumerili", "Pomatomus saltatrix", "Lichia amia")
abbr_chr <- as.character(mean$Species)
full_chr <- as.character(full)
legend("topright",
       inset = c(-0.36, 0),
       legend = rev(
         Map(function(a, f)
           bquote(.(a) ~ "-" ~ italic(.(f))),
           abbr_chr,
           full_chr)),
       bty = "n",
       xpd = TRUE,
       cex = 0.75)
```

```{r}
#Number of individuals by species
num_transmitters <- aggregate(Transmitter ~ Species, data = week_kuds, FUN = function(x) length(unique(x)))

barplot(num_transmitters$Transmitter, 
        names.arg = num_transmitters$Species, 
        xlab = "Species", 
        ylab = "Number of Individuals",
        col = "deepskyblue", 
        border = "black",
        ylim= c(0,180),
        yaxp = c(0, 180, 3),
        cex.names = 0.8,
        las = 2)
text(x = barplot(num_transmitters$Transmitter, plot = FALSE), 
     y = num_transmitters$Transmitter, 
     label = num_transmitters$Transmitter, 
     pos = 3, 
     cex = 0.7)

#Number of observations by species
num_observations <- table(week_kuds$File, week_kuds$Species)
num_observations <- colSums(num_observations)

barplot(num_observations, 
        xlab = "Species", 
        ylab = "Number of Observations", 
        col= c("deepskyblue"),
        border = "black",
        ylim = c(0, 4000),
        yaxp = c(0, 4000, 8),
        cex.names = 0.8,
        las = 2)
```

```{r}
#Spatial visualisation of the numerical biological parameters
ggplot(week_kuds, aes(LengthStd, KUD95, color=Species)) +
         geom_point() +
  labs(title = "KUD95 ~ LengthStd by Species")

ggplot(week_kuds, aes(BodyMassStd, KUD95, color=Species)) +
         geom_point() +
  labs(title = "KUD95 ~ BodyMassStd by Species")

ggplot(week_kuds, aes(Vulnerability, KUD95, color=Species)) +
         geom_point() +
  labs(title = "KUD95 ~ Vulnerability by Species")

ggplot(week_kuds, aes(Longevity, KUD95, color=Species)) +
         geom_point() +
  labs(title = "KUD95 ~ Longevity by Species")

ggplot(week_kuds, aes(Troph, KUD95, color=Species)) +
         geom_point() +
  labs(title = "KUD95 ~ Troph by Species")
```

### Linear Model Assumptions

## Linearity Assumption

# KUD95
```{r}
#Correlation between response and numeric explanatory variables
plot(week_kuds$LengthStd, week_kuds$KUD95, xlab = "Length Std", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of Length Std")
plot(week_kuds$BodyMassStd, week_kuds$KUD95, xlab = "Body Mass Std", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of Body Mass Std")
plot(week_kuds$Longevity, week_kuds$KUD95, xlab = "Longevity", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of Longevity")
plot(week_kuds$Vulnerability, week_kuds$KUD95, xlab = "Vulnerability", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of Vulnerability")
plot(week_kuds$Troph, week_kuds$KUD95, xlab = "Troph", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of Troph")
plot(week_kuds$ReceiverDensity, week_kuds$KUD95, xlab = "Receiver Density", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of Receiver Density")
plot(week_kuds$MonitArea_km2, week_kuds$KUD95, xlab = "MonitArea (km2)", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of Monitored Area")
plot(week_kuds$MCP_km2, week_kuds$KUD95, xlab = "MCP (km2)", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of MCP")
plot(week_kuds$NReceivers, week_kuds$KUD95, xlab = "N Receivers", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of N Receivers")
plot(week_kuds$MaxDistReceivers, week_kuds$KUD95, xlab = "Max. Dist. Receivers", ylab="KUD95", col="deepskyblue", main = "KUD95 as a function of Max. Dist. Receivers")

cor(week_kuds$Length_cm, week_kuds$KUD95)
cor(week_kuds$BodyMass, week_kuds$KUD95)
cor(week_kuds$LengthStd, week_kuds$KUD95)
cor(week_kuds$BodyMassStd, week_kuds$KUD95)
cor(week_kuds$Longevity, week_kuds$KUD95)
cor(week_kuds$Vulnerability, week_kuds$KUD95)
cor(week_kuds$Troph, week_kuds$KUD95)
cor(week_kuds$ReceiverDensity, week_kuds$KUD95)
cor(week_kuds$MonitArea_km2, week_kuds$KUD95)
cor(week_kuds$MCP_km2, week_kuds$KUD95)
cor(week_kuds$NReceivers, week_kuds$KUD95)
cor(week_kuds$MaxDistReceivers, week_kuds$KUD95)
```

# KUD50
```{r}
#Correlation between response and numeric explanatory variables
plot(week_kuds$LengthStd, week_kuds$KUD50, xlab = "Length Std", ylab="KUD50", col="green2", main = "KUD50 as a function of LengthStd")
plot(week_kuds$BodyMassStd, week_kuds$KUD50, xlab = "Body Mass Std", ylab="KUD50", col="green2", main = "KUD50 as a function of Body Mass Std")
plot(week_kuds$Longevity, week_kuds$KUD50, xlab = "Longevity", ylab="KUD50", col="green2", main = "KUD50 as a function of Longevity")
plot(week_kuds$Vulnerability, week_kuds$KUD50, xlab = "Vulnerability", ylab="KUD50", col="green2", main = "KUD50 as a function of Vulnerability")
plot(week_kuds$Troph, week_kuds$KUD50, xlab = "Troph", ylab="KUD50", col="green2", main = "KUD50 as a function of Troph")
plot(week_kuds$ReceiverDensity, week_kuds$KUD50, xlab = "Receiver Density", ylab="KUD50", col="green2", main = "KUD50 as a function of Receiver Density")
plot(week_kuds$MonitArea_km2, week_kuds$KUD50, xlab = "MonitArea (km2)", ylab="KUD50", col="green2", main = "KUD50 as a function of Monitored Area")
plot(week_kuds$MCP_km2, week_kuds$KUD50, xlab = "MCP (km2)", ylab="KUD50", col="green2", main = "KUD50 as a function of MCP")
plot(week_kuds$NReceivers, week_kuds$KUD50, xlab = "N Receivers", ylab="KUD50", col="green2", main = "KUD50 as a function of N Receivers")
plot(week_kuds$MaxDistReceivers, week_kuds$KUD50, xlab = "Max. Dist. Receivers", ylab="KUD50", col="green2", main = "KUD50 as a function of Max. Dist. Receivers")


cor(week_kuds$Length_cm, week_kuds$KUD50)
cor(week_kuds$BodyMass, week_kuds$KUD50)
cor(week_kuds$LengthStd, week_kuds$KUD50)
cor(week_kuds$BodyMassStd, week_kuds$KUD50)
cor(week_kuds$Longevity, week_kuds$KUD50)
cor(week_kuds$Vulnerability, week_kuds$KUD50)
cor(week_kuds$Troph, week_kuds$KUD50)
cor(week_kuds$ReceiverDensity, week_kuds$KUD50)
cor(week_kuds$MonitArea_km2, week_kuds$KUD50)
cor(week_kuds$MCP_km2, week_kuds$KUD50)
cor(week_kuds$NReceivers, week_kuds$KUD50)
cor(week_kuds$MaxDistReceivers, week_kuds$KUD50)

cor.data <- data.frame(week_kuds$KUD95, week_kuds$KUD50, week_kuds$LengthStd, week_kuds$BodyMassStd, week_kuds$Longevity, week_kuds$Vulnerability, week_kuds$Troph, week_kuds$ReceiverDensity, week_kuds$MonitArea_km2, week_kuds$MCP_km2, week_kuds$NReceivers, week_kuds$MaxDistReceivers)
names(cor.data) <- c("KUD95", "KUD50", "LengthStd", "BodyMassStd", "Longevity", "Vulnerability", "Troph", "ReceiverDensity", "MonitArea_km2", "MCP_km2", "NReceivers", "MaxDistReceivers")
cortable <- cor(cor.data)
corrplot(cortable, tl.col = "black")
```

It doesn't seem to exist a linear relationship between the response variables and any of the numeric explanatory variables, since the correlation is approximately 0.


To see if makes sense to introduce a certain categorical variable in a linear model, we first need to observe if the different categories of each variable are easily distinguished by their response variable values.
```{r}
#Boxplots of categorical explanatory variables KUD95

par(mfrow = c(2, 3))
boxplot(KUD95 ~ Habitat, data= week_kuds, col="deepskyblue", xlab="Habitat")
title("Boxplot of KUD95 given Habitat")
boxplot(KUD95 ~ Migration, data= week_kuds, col="deepskyblue", xlab = "Migration") 
title("Boxplot of KUD95 given Migration")

leveneTest(KUD95 ~ Habitat, data = week_kuds)
leveneTest(KUD95 ~ Migration, data = week_kuds)

#Boxplots of categorical explanatory variables KUD50

par(mfrow = c(1, 2))
boxplot(KUD50 ~ Habitat, data= week_kuds, col="green2", xlab="Habitat")
title("Boxplot of KUD50 given Habitat")
boxplot(KUD50 ~ Migration, data= week_kuds, col="green2", xlab = "Migration") 
title("Boxplot of KUD50 given Migration")

leveneTest(KUD50 ~ Habitat, data = week_kuds)
leveneTest(KUD50 ~ Migration, data = week_kuds)
kruskal.test(KUD95 ~ Habitat, data = week_kuds)
kruskal.test(KUD95 ~ Migration, data = week_kuds)
```

We made boxplots to observe if the values of each category were different from each other, but the only thing we could see was that there was overlapping in the values of the boxes and a lot of outliers. Also, through the levene test we could observe that the variance between groups in each categorical variable is heterogeneous. 

The data violates the Linearity Assumption.

## Multicolinearity Inspection (Correlation Between Parameters)
```{r}
#Biological traits correlation
cor(week_kuds$Length_cm, week_kuds$BodyMass)
cor(week_kuds$Length_cm, week_kuds$LengthStd)
cor(week_kuds$Length_cm, week_kuds$BodyMassStd)
cor(week_kuds$Length_cm, week_kuds$Troph)
cor(week_kuds$Length_cm, week_kuds$Vulnerability)
cor(week_kuds$Length_cm, week_kuds$Longevity)
cor(week_kuds$BodyMass, week_kuds$LengthStd)
cor(week_kuds$BodyMass, week_kuds$BodyMassStd)
cor(week_kuds$BodyMass, week_kuds$Troph)
cor(week_kuds$BodyMass, week_kuds$Vulnerability)
cor(week_kuds$BodyMass, week_kuds$Longevity)
cor(week_kuds$LengthStd, week_kuds$BodyMassStd)
cor(week_kuds$LengthStd, week_kuds$Longevity)
cor(week_kuds$LengthStd, week_kuds$Vulnerability)
cor(week_kuds$LengthStd, week_kuds$Troph)
cor(week_kuds$BodyMassStd, week_kuds$Longevity)
cor(week_kuds$BodyMassStd, week_kuds$Vulnerability)
cor(week_kuds$BodyMassStd, week_kuds$Troph)
cor(week_kuds$Longevity, week_kuds$Vulnerability)
cor(week_kuds$Longevity, week_kuds$Troph)
cor(week_kuds$Vulnerability, week_kuds$Troph)

plot(correlation(week_kuds[,c(11, 13, 14, 15, 16)], method = "pearson"), main="Linear correlation between biological traits")
cor.matrix1 <- data.frame(week_kuds$LengthStd, week_kuds$BodyMassStd, week_kuds$Longevity, week_kuds$Vulnerability, week_kuds$Troph)
names(cor.matrix1) <- c("LengthStd", "BodyMassStd", "Longevity", "Vulnerability", "Troph")
cortable1 <- cor(cor.matrix1)
corrplot(cortable1, tl.col = "black")

#Experimental design parameters correlation
cor(week_kuds$ReceiverDensity, week_kuds$MonitArea_km2)
cor(week_kuds$ReceiverDensity, week_kuds$MCP_km2)
cor(week_kuds$ReceiverDensity, week_kuds$NReceivers)
cor(week_kuds$ReceiverDensity, week_kuds$MaxDistReceivers)
cor(week_kuds$MonitArea_km2, week_kuds$MCP_km2)
cor(week_kuds$MonitArea_km2, week_kuds$NReceivers)
cor(week_kuds$MonitArea_km2, week_kuds$MaxDistReceivers)
cor(week_kuds$MCP_km2, week_kuds$NReceivers)
cor(week_kuds$MCP_km2, week_kuds$MaxDistReceivers)
cor(week_kuds$NReceivers, week_kuds$MaxDistReceivers)

plot(correlation(week_kuds[,c(17, 18, 19, 20, 21)]), main="Linear correlation between experimental design parameters")
cor.data2 <- data.frame(week_kuds$ReceiverDensity, week_kuds$MonitArea_km2, week_kuds$MCP_km2, week_kuds$NReceivers, week_kuds$MaxDistReceivers)
names(cor.data2) <- c("ReceiverDensity", "MonitArea_km2", "MCP_km2", "NReceivers", "MaxDistReceivers")
cortable2 <- cor(cor.data2)
corrplot(cortable2, tl.col = "black")
```

We can observe that the only biological traits with a relative high linear correlation are the length standardised (LenghtStd) with the body mass standardised (BodyMassStd), with 74,7%. For the experimental design parameters, the monitored area has a relative high linear correlation with the number of receivers (NReceivers) and with the maximum distance between receivers (MaxDistReceivers), with 98,4% and 74,4% respectively. In future possible linear model analysis we may choose to include only one of these variables to avoid errors in coefficient estimates.

## Normality and Homocedasticity Assumptions

We tested the full linear model and checked the assumptions.

# KUD95
```{r}
#Full model assumption checking
lmtotal <- lm(KUD95 ~ LengthStd + BodyMassStd + Longevity + Vulnerability + Troph + Habitat + Migration + ReceiverDensity + MonitArea_km2, data = week_kuds)
summary(lmtotal)

#Normality assumption
qqnorm(lmtotal$residuals)
qqline(lmtotal$residuals, col = 2)

#Homocedascity assumption
plot(lmtotal$fitted.values, lmtotal$residuals)
abline(h=0, col="red")
```

# KUD50
```{r}
#Full model assumption checking
lmtotal <- lm(KUD50 ~ LengthStd + BodyMassStd + Longevity + Vulnerability + Troph + Habitat + Migration + ReceiverDensity + MonitArea_km2, data = week_kuds)
summary(lmtotal)

#Normality assumption
qqnorm(lmtotal$residuals)
qqline(lmtotal$residuals, col = 2)

#Homocedascity assumption
plot(lmtotal$fitted.values, lmtotal$residuals)
abline(h=0, col="red")
```

As expected, the linear model do not fit well to our data, since the assumptions are not met. 



We tried to fit different types of models. We modeled the response against each predictor variable separately. We then compare the AIC of each model to see which one fitted better.

```{r}
#Modeling (each variable separately)

#Length Standardised
lm_length <- lm(KUD95 ~ LengthStd, data=week_kuds)
glm_length <- glm(KUD95 ~ LengthStd, data=week_kuds, family=Gamma(link="log"))
gam_length <- gam(KUD95 ~ LengthStd, data=week_kuds, family=Gamma(link="log"))
glmmF_length <- glmmTMB(KUD95 ~ LengthStd + (1|File), data=week_kuds, family=Gamma(link="log"))
glmmT_length <- glmmTMB(KUD95 ~ LengthStd + (1|Transmitter), data=week_kuds, family=Gamma(link="log")) #control = glmerControl(optimizer = "bobyqa")
glmmS_length <- glmmTMB(KUD95 ~ LengthStd + (1|Species), data=week_kuds, family=Gamma(link="log"))
gammF_length <- gamm(KUD95 ~ s(LengthStd), random=list(File=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4F_length <- gamm4(KUD95 ~ s(LengthStd), random = ~(1|File), data = week_kuds, family=Gamma(link="log"))
gammT_length <- gamm(KUD95 ~ s(LengthStd), random=list(Transmitter=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4T_length <- gamm4(KUD95 ~ s(LengthStd), random = ~(1|Transmitter), data = week_kuds, family=Gamma(link="log"))
gammS_length <- gamm(KUD95 ~ s(LengthStd), random=list(Species=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4S_length <- gamm4(KUD95 ~ s(LengthStd), random = ~(1|Species), data = week_kuds, family=Gamma(link="log"))


AIC(lm_length, glm_length, gam_length, glmmF_length, glmmT_length, glmmS_length)
summary(gammF_length$lme)
summary(gammT_length$lme)
summary(gammS_length$lme)


#BodyMass Standardised
lm_bodymass <- lm(KUD95 ~ BodyMassStd, data=week_kuds)
glm_bodymass <- glm(KUD95 ~ BodyMassStd, data=week_kuds, family=Gamma(link="log"))
gam_bodymass <- gam(KUD95 ~ BodyMassStd, data=week_kuds, family=Gamma(link="log"))
glmmF_bodymass <- glmmTMB(KUD95 ~ BodyMassStd + (1|File), data=week_kuds, family=Gamma(link="log"))
glmmT_bodymass <- glmmTMB(KUD95 ~ BodyMassStd + (1|Transmitter), data=week_kuds, family=Gamma(link="log"))
glmmS_bodymass <- glmmTMB(KUD95 ~ BodyMassStd + (1|Species), data=week_kuds, family=Gamma(link="log"))
gammF_bodymass <- gamm(KUD95 ~ s(BodyMassStd), random=list(File=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4F_bodymass <- gamm4(KUD95 ~ s(BodyMassStd), random = ~(1|File), data = week_kuds, family=Gamma(link="log"))
gammT_bodymass <- gamm(KUD95 ~ s(BodyMassStd), random=list(Transmitter=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4T_bodymass <- gamm4(KUD95 ~ s(BodyMassStd), random = ~(1|Transmitter), data = week_kuds, family=Gamma(link="log"))
gammS_bodymass <- gamm(KUD95 ~ s(BodyMassStd), random=list(Species=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4S_bodymass <- gamm4(KUD95 ~ s(BodyMassStd), random = ~(1|Species), data = week_kuds, family=Gamma(link="log"))


AIC(lm_bodymass, glm_bodymass, gam_bodymass, glmmF_bodymass, glmmT_bodymass, glmmS_bodymass)
summary(gammF_bodymass$lme)
summary(gammT_bodymass$lme)
summary(gammS_bodymass$lme)


#Longevity
lm_longevity <- lm(KUD95 ~ Longevity, data=week_kuds)
glm_longevity <- glm(KUD95 ~ Longevity, data=week_kuds, family=Gamma(link="log"))
gam_longevity <- gam(KUD95 ~ Longevity, data=week_kuds, family=Gamma(link="log"))
glmmF_longevity <- glmmTMB(KUD95 ~ Longevity + (1|File), data=week_kuds, family=Gamma(link="log"))
glmmT_longevity <- glmmTMB(KUD95 ~ Longevity + (1|Transmitter), data=week_kuds, family=Gamma(link="log"))
glmmS_longevity <- glmmTMB(KUD95 ~ Longevity + (1|Species), data=week_kuds, family=Gamma(link="log"))
gammF_longevity <- gamm(KUD95 ~ s(Longevity), random=list(File=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4F_longevity <- gamm4(KUD95 ~ s(Longevity), random = ~(1|File), data = week_kuds, family=Gamma(link="log"))
gammT_longevity <- gamm(KUD95 ~ s(Longevity), random=list(Transmitter=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4T_longevity <- gamm4(KUD95 ~ s(Longevity), random = ~(1|Transmitter), data = week_kuds, family=Gamma(link="log"))
gammS_longevity <- gamm(KUD95 ~ s(Longevity), random=list(Species=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4S_longevity <- gamm4(KUD95 ~ s(Longevity), random = ~(1|Species), data = week_kuds, family=Gamma(link="log"))

AIC(lm_longevity, glm_longevity, gam_longevity, glmmF_longevity, glmmT_longevity, glmmS_longevity)
summary(gammF_longevity$lme)
summary(gammT_longevity$lme)
summary(gammS_longevity$lme)


#Vulnerability
lm_vulnerability <- lm(KUD95 ~ Vulnerability, data=week_kuds)
glm_vulnerability <- glm(KUD95 ~ Vulnerability, data=week_kuds, family=Gamma(link="log"))
gam_vulnerability <- gam(KUD95 ~ Vulnerability, data=week_kuds, family=Gamma(link="log"))
glmmF_vulnerability <- glmmTMB(KUD95 ~ Vulnerability + (1|File), data=week_kuds, family=Gamma(link="log"))
glmmT_vulnerability <- glmmTMB(KUD95 ~ Vulnerability + (1|Transmitter), data=week_kuds, family=Gamma(link="log"))
glmmS_vulnerability <- glmmTMB(KUD95 ~ Vulnerability + (1|Species), data=week_kuds, family=Gamma(link="log"))
gammF_vulnerability <- gamm(KUD95 ~ s(Vulnerability), random=list(File=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4F_vulnerability <- gamm4(KUD95 ~ s(Vulnerability), random = ~(1|File), data = week_kuds, family=Gamma(link="log"))
gammT_vulnerability <- gamm(KUD95 ~ s(Vulnerability), random=list(Transmitter=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4T_vulnerability <- gamm4(KUD95 ~ s(Vulnerability), random = ~(1|Transmitter), data = week_kuds, family=Gamma(link="log"))
gammS_vulnerability <- gamm(KUD95 ~ s(Vulnerability), random=list(Species=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4S_vulnerability <- gamm4(KUD95 ~ s(Vulnerability), random = ~(1|Species), data = week_kuds, family=Gamma(link="log"))

AIC(lm_vulnerability, glm_vulnerability, gam_vulnerability, glmmF_vulnerability, glmmT_vulnerability, glmmS_vulnerability)
summary(gammF_vulnerability$lme)
summary(gammT_vulnerability$lme)
summary(gammS_vulnerability$lme)


#Troph
lm_troph <- lm(KUD95 ~ Troph, data=week_kuds)
glm_troph <- glm(KUD95 ~ Troph, data=week_kuds, family=Gamma(link="log"))
gam_troph <- gam(KUD95 ~ Troph, data=week_kuds, family=Gamma(link="log"))
glmmF_troph <- glmmTMB(KUD95 ~ Troph + (1|File), data=week_kuds, family=Gamma(link="log"))
glmmT_troph <- glmmTMB(KUD95 ~ Troph + (1|Transmitter), data=week_kuds, family=Gamma(link="log"))
glmmS_troph <- glmmTMB(KUD95 ~ Troph + (1|Species), data=week_kuds, family=Gamma(link="log"))
gammF_troph <- gamm(KUD95 ~ s(Troph), random=list(File=~1), data= week_kuds, family=Gamma(link="log"))   ##Preference for gmcv package since gamm4 uses glm methods
#gamm4F_troph <- gamm4(KUD95 ~ s(Troph), random = ~(1|File), data = week_kuds, family=Gamma(link="log"))
gammT_troph <- gamm(KUD95 ~ s(Troph), random=list(Transmitter=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4T_troph <- gamm4(KUD95 ~ s(Troph), random = ~(1|Transmitter), data = week_kuds, family=Gamma(link="log"))
gammS_troph <- gamm(KUD95 ~ s(Troph), random=list(Species=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4S_troph <- gamm4(KUD95 ~ s(Troph), random = ~(1|Species), data = week_kuds, family=Gamma(link="log"))

AIC(lm_troph, glm_troph, gam_troph, glmmF_troph, glmmT_troph, glmmS_troph)
summary(gammF_troph$lme)
summary(gammT_troph$lme)
summary(gammS_troph$lme)


#Habitat
lm_habitat <- lm(KUD95 ~ Habitat, data=week_kuds)
glm_habitat <- glm(KUD95 ~ Habitat, data=week_kuds, family=Gamma(link="log"))
gam_habitat <- gam(KUD95 ~ Habitat, data=week_kuds, family=Gamma(link="log"))
glmmF_habitat <- glmmTMB(KUD95 ~ Habitat + (1|File), data=week_kuds, family=Gamma(link="log"))
glmmT_habitat <- glmmTMB(KUD95 ~ Habitat + (1|Transmitter), data=week_kuds, family=Gamma(link="log"))
glmmS_habitat <- glmmTMB(KUD95 ~ Habitat + (1|Species), data=week_kuds, family=Gamma(link="log"))
gammF_habitat <- gamm(KUD95 ~ Habitat, random=list(File=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4F_habitat <- gamm4(KUD95 ~ Habitat, random = ~(1|File), data = week_kuds, family=Gamma(link="log"))
gammT_habitat <- gamm(KUD95 ~ Habitat, random=list(Transmitter=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4T_habitat <- gamm4(KUD95 ~ Habitat, random = ~(1|Transmitter), data = week_kuds, family=Gamma(link="log"))
gammS_habitat <- gamm(KUD95 ~ Habitat, random=list(Species=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4S_habitat <- gamm4(KUD95 ~ Habitat, random = ~(1|Species), data = week_kuds, family=Gamma(link="log"))

AIC(lm_habitat, glm_habitat, gam_habitat, glmmF_habitat, glmmT_habitat, glmmS_habitat)
summary(gammF_habitat$lme)
summary(gammT_habitat$lme)
summary(gammS_habitat$lme)


#Migration
lm_migration <- lm(KUD95 ~ Migration, data=week_kuds)
glm_migration <- glm(KUD95 ~ Migration, data=week_kuds, family=Gamma(link="log"))
gam_migration <- gam(KUD95 ~ Migration, data=week_kuds, family=Gamma(link="log"))
glmmF_migration <- glmmTMB(KUD95 ~ Migration + (1|File), data=week_kuds, family=Gamma(link="log"))
glmmT_migration <- glmmTMB(KUD95 ~ Migration + (1|Transmitter), data=week_kuds, family=Gamma(link="log"))
glmmS_migration <- glmmTMB(KUD95 ~ Migration + (1|Species), data=week_kuds, family=Gamma(link="log"))
gammF_migration <- gamm(KUD95 ~ Migration, random=list(File=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4F_migration <- gamm4(KUD95 ~ Migration, random = ~(1|File), data = week_kuds, family=Gamma(link="log"))
gammT_migration <- gamm(KUD95 ~ Migration, random=list(Transmitter=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4T_migration <- gamm4(KUD95 ~ Migration, random = ~(1|Transmitter), data = week_kuds, family=Gamma(link="log"))
gammS_migration <- gamm(KUD95 ~ Migration, random=list(Species=~1), data= week_kuds, family=Gamma(link="log"))   #Preference for gmcv package since gamm4 uses glm methods
#gamm4S_migration <- gamm4(KUD95 ~ Migration, random = ~(1|Species), data = week_kuds, family=Gamma(link="log"))

AIC(lm_migration, glm_migration, gam_migration, glmmF_migration, glmmT_migration, glmmS_migration)
summary(gammF_migration$lme)
summary(gammT_migration$lme)
summary(gammS_migration$lme)
```

Observing the AIC values of all models adjusted, we can see that the one that better fits the data is the GLMM, with the Transmitter as random effect. Looking for all the GLMM adjusted, with higher attention to the variables used as random effects, we see that the models adjusted with Transmitter are better than the ones adjusted with File, and lastly Species. However, we decided to include Species and Transmitter as nested random effects to account for potential pseudoreplication and non-independence of observations from the same species and individuals, and it presents the lowest AIC.



Having chosen the best variables to include in the model as random effects, we first evaluated the significance of each biological trait in explaining the KUDs alone.
```{r}
#testing all variables separately using GLMM with File and Transmitter as random effects
options(scipen = 999)

#Each trait with all the design variables included (KUD95)
summary(glmmTMB(KUD95 ~ Length_cm + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~ BodyMass + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~ LengthStd + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~ BodyMassStd + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~ Longevity + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~ Vulnerability + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~ Troph + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~ Habitat + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers +  (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~ Spawn + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers +  (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD95 ~  Migration + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers +  (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))

#Each trait with all the design variables included (KUD50)
summary(glmmTMB(KUD50 ~ Length_cm + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~ BodyMass + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~ LengthStd + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~ BodyMassStd + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~ Longevity + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~ Vulnerability + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~ Troph + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~ Habitat + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers +  (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~ Spawn + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers +  (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
summary(glmmTMB(KUD50 ~  Migration + MonitArea_km2 + ReceiverDensity + MCP_km2 + NReceivers + MaxDistReceivers +  (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")))
```

We found that trophic level, habitat, spawning season and migration are related with the KUDs.

```{r}
#Code to generate Figure 3-B

m_troph <- glmmTMB(KUD95 ~ Troph + MonitArea_km2 + ReceiverDensity + MCP_km2 +
                     NReceivers + MaxDistReceivers + (1|Species/Transmitter),
                   data = week_kuds, family = Gamma(link="log"))

m_habitat <- glmmTMB(KUD95 ~ Habitat + MonitArea_km2 + ReceiverDensity + MCP_km2 +
                       NReceivers + MaxDistReceivers + (1|Transmitter) + (1|Species),
                     data = week_kuds, family = Gamma(link="log"))

m_spawn <- glmmTMB(KUD95 ~ Spawn + MonitArea_km2 + ReceiverDensity + MCP_km2 +
                     NReceivers + MaxDistReceivers + (1|Transmitter) + (1|Species),
                   data = week_kuds, family = Gamma(link="log"))

m_migr <- glmmTMB(KUD95 ~ Migration + MonitArea_km2 + ReceiverDensity + MCP_km2 +
                    NReceivers + MaxDistReceivers + (1|Transmitter) + (1|Species),
                  data = week_kuds, family = Gamma(link="log"))

m_troph50 <- glmmTMB(KUD50 ~ Troph + MonitArea_km2 + ReceiverDensity + MCP_km2 +
                       NReceivers + MaxDistReceivers + (1|Species/Transmitter),
                     data = week_kuds, family = Gamma(link="log"))

m_habitat50 <- glmmTMB(KUD50 ~ Habitat + MonitArea_km2 + ReceiverDensity + MCP_km2 +
                         NReceivers + MaxDistReceivers + (1|Transmitter) + (1|Species),
                       data = week_kuds, family = Gamma(link="log"))

m_spawn50 <- glmmTMB(KUD50 ~ Spawn + MonitArea_km2 + ReceiverDensity + MCP_km2 +
                       NReceivers + MaxDistReceivers + (1|Transmitter) + (1|Species),
                     data = week_kuds, family = Gamma(link="log"))

m_migr50 <- glmmTMB(KUD50 ~ Migration + MonitArea_km2 + ReceiverDensity + MCP_km2 +
                      NReceivers + MaxDistReceivers + (1|Transmitter) + (1|Species),
                    data = week_kuds, family = Gamma(link="log"))

g_troph <- ggpredict(m_troph, terms = "Troph")
g_habitat <- ggpredict(m_habitat, terms = "Habitat")
g_spawn <- ggpredict(m_spawn, terms = "Spawn")
g_migr <- ggpredict(m_migr, terms = "Migration")
g_troph50   <- ggpredict(m_troph50, terms = "Troph")
g_habitat50 <- ggpredict(m_habitat50, terms = "Habitat")
g_spawn50   <- ggpredict(m_spawn50, terms = "Spawn")
g_migr50    <- ggpredict(m_migr50, terms = "Migration")

p_troph <- ggplot(g_troph, aes(x, predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "deepskyblue4") +
  geom_line(size = 1.3, color = "deepskyblue4") +
  labs(x= "Trophic Level", y = "Trophic Level") +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.y = element_text(margin = margin(r = 15))  # aumenta espaço à direita do rótulo
  )

p_habitat <- ggplot(g_habitat, aes(x, predicted)) +
  geom_point(size = 3, color = c("deepskyblue", "deepskyblue3","deepskyblue4")) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3, color = c("deepskyblue", "deepskyblue3","deepskyblue4"), linewidth = 1.3) +
  labs(
    title = "Total range",
    subtitle = expression("Predicted weekly KUD95 (km"^2*")"),
    x = " ",
    y = "Habitat"
  ) +
  theme_minimal(base_size = 14) +
   theme(
     axis.title.y = element_text(margin = margin(r = 15)),
    plot.title = element_text(color = "deepskyblue4", size = 15, face = "bold", hjust = 0.5, margin = margin(b = 1)),
    plot.subtitle = element_text(color = "deepskyblue4", size = 10, face = "bold", hjust = 0.5)
  )

p_spawn <- ggplot(g_spawn, aes(x, predicted)) +
  geom_point(size = 3, color = c("deepskyblue","deepskyblue4")) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = c("deepskyblue","deepskyblue4"), linewidth = 1.3) +
  labs(x = " ", y = "Spawning Period") +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.y = element_text(margin = margin(r = 15))
  )

p_migr <- ggplot(g_migr, aes(x, predicted)) +
  geom_point(size = 3, color = c("deepskyblue","deepskyblue4")) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = c("deepskyblue","deepskyblue4"), linewidth = 1.3) +
  labs(x = " ", y = "Migration") +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.y = element_text(margin = margin(r = 15))
  )

p_troph50 <- ggplot(g_troph50, aes(x, predicted)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2, fill = "forestgreen") +
  geom_line(size = 1.3, color = "forestgreen") +
  labs(x = "Trophic Level", y = " ") +
  theme_minimal(base_size = 14)

p_habitat50 <- ggplot(g_habitat50, aes(x, predicted)) +
  geom_point(size = 3, color = c("green", "limegreen","forestgreen")) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3, color = c("green", "limegreen","forestgreen"), linewidth = 1.3) +
  labs(
    title = "Total range",
    subtitle = expression("Predicted weekly KUD50 (km"^2*")"),
    x = " ",
    y = " "
  ) +
  theme_minimal(base_size = 12) +
   theme(
    plot.title = element_text(color = "forestgreen", size = 15, face = "bold", hjust = 0.5, margin = margin(b = 1)),
    plot.subtitle = element_text(color = "forestgreen", size = 10, face = "bold", hjust = 0.5)
  )

p_migr50 <- ggplot(g_migr50, aes(x, predicted)) +
  geom_point(size = 3, color = c("green", "forestgreen")) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = c("green", "forestgreen"), linewidth = 1.3) +
  labs(x = " ", y = " ") +
  theme_minimal(base_size = 12)

p_spawn50 <- ggplot(g_spawn50, aes(x, predicted)) +
  geom_point(size = 3, color = c("green", "forestgreen")) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = c("green", "forestgreen"), linewidth = 1.3) +
  labs(x = " ", y = " ") +
  theme_minimal(base_size = 12)


(p_habitat | p_migr | p_troph | p_spawn) /
(p_habitat50 | p_migr50 | p_troph50 | p_spawn50)

(p_habitat | p_habitat50) /
(p_migr | p_migr50) /
(p_troph | p_troph50) /
(p_spawn | p_spawn50)

```

However, we investigated if the models could reach a better fit by including more variables. We did that be performing Backward Elimination.

# KUD95 Backward Elimination

```{r}
Final1 <- glmmTMB(KUD95 ~ Length_cm + BodyMass + LengthStd + BodyMassStd + Longevity + Vulnerability + Troph + Habitat + Migration + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final1)

Final2 <- glmmTMB(KUD95 ~ Length_cm + LengthStd + BodyMassStd + Longevity + Vulnerability + Troph + Habitat + Migration + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final2)

Final3 <- glmmTMB(KUD95 ~ Length_cm + LengthStd + BodyMassStd + Longevity + Troph + Habitat + Migration + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final3)

Final4 <- glmmTMB(KUD95 ~ Length_cm + LengthStd + BodyMassStd + Longevity + Troph + Habitat + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final4)

Final5 <- glmmTMB(KUD95 ~ Length_cm + LengthStd + BodyMassStd + Longevity + Troph + Habitat + Spawn + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final5)

Final6 <- glmmTMB(KUD95 ~ Length_cm + LengthStd + BodyMassStd + Troph + Habitat + Spawn + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final6)

Final7 <- glmmTMB(KUD95 ~ Length_cm + LengthStd + Troph + Habitat + Spawn + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final7)

Final8 <- glmmTMB(KUD95 ~ Length_cm + LengthStd + Habitat + Spawn + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final8)

Final9 <- glmmTMB(KUD95 ~ LengthStd + Habitat + Spawn + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final9)

Final10 <- glmmTMB(KUD95 ~ Habitat + Spawn + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final10)

anova(Final1, Final10) #as there is no evidence that this two models are statistically different from each other, we must choose the simplest one, the one with less variables (Final7)

ranef(Final10) #deviation from intercept for each Transmitter and File group

confint(Final10)


#evaluate if the final model is better than the ones including only one biological trait
anova(glmmTMB(KUD95 ~ Habitat + MonitArea_km2 + MCP_km2 + ReceiverDensity + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")), Final10)
anova(glmmTMB(KUD95 ~ Spawn + MonitArea_km2 + MCP_km2 + ReceiverDensity + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")), Final10)
```

We reached the best fitted model (Final10). Now we must analyse the residuals to see how they behave.


```{r}
#Test residuals for the best model (goodness of fit)
testDispersion(Final10)  #plot with normality, dispersion and outliers

simulationOutput <- simulateResiduals(fittedModel = Final10, plot = F) #dispersion test
testDispersion(simulationOutput) #dispersion test

plot(simulationOutput) #residual analysis
plotQQunif(simulationOutput) #Q-Q plot (normality checking)
plotResiduals(simulationOutput) #residual vs. predicted (homoscedasticity checking)
testOutliers(simulationOutput) #outliers checking

#Simulations from the model
getObservedResponse(Final10)  #response used to fit the model
getFitted(Final10)  #predictions of the model for all points
# Gerar 100 simulações
set.seed(123)  # para reprodutibilidade
x <- getSimulations(Final10, nsim = 100, type = "refit")  #extract simulations from the model
sim_matrix <- do.call(cbind, x)
sim_mean <- rowMeans(sim_matrix)
df <- data.frame(
  KUD95_real = week_kuds$KUD95,
  KUD95_sim_mean = sim_mean,
  Habitat = week_kuds$Habitat,
  Spawn = week_kuds$Spawn
)

p_real <- ggplot(df, aes(x = Habitat, y = KUD95_real)) +
  geom_boxplot(fill = "gray") +
  scale_y_continuous(limits = c(0, 15)) +
  xlab("Habitat") +
  ylab("Real KUD95") +
  theme_minimal()
p_sim <- ggplot(df, aes(x = Habitat, y = KUD95_sim_mean)) +
  geom_boxplot(fill = "deepskyblue3") +
  scale_y_continuous(limits = c(0, 15)) +
  xlab("Habitat") +
  ylab("Mean simulated KUD95") +
  theme_minimal()

grid.arrange(p_real, p_sim, ncol = 2)

p_real <- ggplot(df, aes(x = Spawn, y = KUD95_real)) +
  geom_boxplot(fill = "gray") +
  scale_y_continuous(limits = c(0, 15)) +
  xlab("Spawning season") +
  ylab("Real KUD95") +
  theme_minimal()
p_sim <- ggplot(df, aes(x = Spawn, y = KUD95_sim_mean)) +
  geom_boxplot(fill = "deepskyblue3") +
  scale_y_continuous(limits = c(0, 15)) +
  xlab("Spawning season") +
  ylab("Mean simulated KUD95") +
  theme_minimal()

grid.arrange(p_real, p_sim, ncol = 2)
```

The normality and homoscedasticity assumptions aren't met. However, given the large dataset, this violation may not be a big problem. To investigate this, we simulated the response values and compared with the real ones. After observing the results, we concluded that the patterns were similar and the model can be said to be well adjusted.


# KUD50 Backward Elimination

```{r}
Final1.1 <- glmmTMB(KUD50 ~ Length_cm + BodyMass + LengthStd + BodyMassStd + Longevity + Vulnerability + Troph + Habitat + Migration + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final1.1)

Final2.1 <- glmmTMB(KUD50 ~ Length_cm + BodyMass + LengthStd + Longevity + Vulnerability + Troph + Habitat + Migration + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final2.1)

Final3.1 <- glmmTMB(KUD50 ~ Length_cm + BodyMass + LengthStd + Longevity + Troph + Habitat + Migration + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final3.1)

Final4.1 <- glmmTMB(KUD50 ~ Length_cm + BodyMass + LengthStd + Longevity + Troph + Habitat + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final4.1)

Final5.1 <- glmmTMB(KUD50 ~ Length_cm + LengthStd + Longevity + Troph + Habitat + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final5.1)

Final6.1 <- glmmTMB(KUD50 ~ Length_cm + LengthStd + Troph + Habitat + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final6.1)

Final7.1 <- glmmTMB(KUD50 ~ Length_cm + LengthStd + Habitat + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final7.1)

Final8.1 <- glmmTMB(KUD50 ~ LengthStd + Habitat + Spawn + ReceiverDensity + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final8.1)

Final9.1 <- glmmTMB(KUD50 ~ LengthStd + Habitat + Spawn + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final9.1)

Final10.1 <- glmmTMB(KUD50 ~ Habitat + Spawn + MonitArea_km2 + MCP_km2 + (1|Species/Transmitter), data = week_kuds, family = Gamma(link="log"))
summary(Final10.1)

anova(Final1.1, Final8.1) #as there is no evidence that this two models are statistically different from each other, we must choose the simplest one, the one with less variables (Final7)

ranef(Final8.1) #deviation from intercept for each Transmitter and File group

confint(Final8.1)

#evaluate if the final model is better than the ones including only one biological trait
anova(glmmTMB(KUD50 ~ Habitat + MonitArea_km2 + MCP_km2 + ReceiverDensity + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")),Final10.1)
anova(glmmTMB(KUD50 ~ Spawn + MonitArea_km2 + MCP_km2 + ReceiverDensity + (1|Species/Transmitter), data=week_kuds, family=Gamma(link="log")), Final10.1)
```

We reached the best fitted model. Now we must analyse the residuals to see how they behave.


```{r}
#Test residuals for the best model (goodness of fit)
testDispersion(Final10.1)  #plot with normality, dispersion and outliers

simulationOutput <- simulateResiduals(fittedModel = Final10.1, plot = F) #dispersion test
testDispersion(simulationOutput) #dispersion test

plot(simulationOutput) #residual analysis
plotQQunif(simulationOutput) #Q-Q plot (normality checking)
plotResiduals(simulationOutput) #residual vs. predicted (homoscedasticity checking)
testOutliers(simulationOutput) #outliers checking

#Simulations from the model
getObservedResponse(Final10.1)  #response used to fit the model
getFitted(Final10.1)  #predictions of the model for all points
# Gerar 100 simulações
set.seed(123)  # para reprodutibilidade
x <- getSimulations(Final10.1, nsim = 100, type = "refit")  #extract simulations from the model
sim_matrix <- do.call(cbind, x)
sim_mean <- rowMeans(sim_matrix)
df <- data.frame(
  KUD50_real = week_kuds$KUD50,
  KUD50_sim_mean = sim_mean,
  Habitat = week_kuds$Habitat,
  Spawn = week_kuds$Spawn
)

p_real <- ggplot(df, aes(x = Habitat, y = KUD50_real)) +
  geom_boxplot(fill = "gray") +
  scale_y_continuous(limits = c(0, 15)) +
  xlab("Habitat") +
  ylab("Real KUD50") +
  theme_minimal()
p_sim <- ggplot(df, aes(x = Habitat, y = KUD50_sim_mean)) +
  geom_boxplot(fill = "green") +
  scale_y_continuous(limits = c(0, 15)) +
  xlab("Habitat") +
  ylab("Mean simulated KUD50") +
  theme_minimal()

grid.arrange(p_real, p_sim, ncol = 2)

p_real <- ggplot(df, aes(x = Spawn, y = KUD50_real)) +
  geom_boxplot(fill = "gray") +
  scale_y_continuous(limits = c(0, 15)) +
  xlab("Spawning season") +
  ylab("Real KUD50") +
  theme_minimal()
p_sim <- ggplot(df, aes(x = Spawn, y = KUD50_sim_mean)) +
  geom_boxplot(fill = "green") +
  scale_y_continuous(limits = c(0, 15)) +
  xlab("Spawning season") +
  ylab("Mean simulated KUD50") +
  theme_minimal()

grid.arrange(p_real, p_sim, ncol = 3)
```

The normality and homoscedasticity assumptions aren't met. However, given the large dataset, this violation may not be a big problem. To investigate this, we simulated the response values and compared with the real ones. After observing the results, we concluded that the patterns were similar and the model can be said to be well adjusted.

